---
title: "A3: SPL Library Checkouts"
output: html_document
editor_options: 
  markdown: 
    wrap: sentence
---

## Library Checkout Trends for author Susan Cain

### Chantria Im

### February 24th, 2023

### Introduction

For my research, I will be analyzing library checkout trends by New York Times bestselling author Susan Cain from a dataset collected by the *City of Seattle Open Data Portal* from January 2012 to January 2023.
Cain is best known for her books *Quiet: The Power of Introverts* (published on January 24, 2012) and *Bittersweet: How Sorrow and Longing Make Us Whole*, (published on April 5, 2022).
As an avid reader, introvert and melancholic individual, I naturally became gravitated to these books to gain a better understanding of myself and transform the way I view life.
I wanted to analyze these trends to draw conclusions about possible reasons why a certain month or year had a spike or drop in library checkouts.

### References

Seattle.(n.d.).
Checkouts by Title.
Retrieved from Seattle Open Data Portal

### Summary Information

Write a summary paragraph of findings that includes the 5 values calculated from your summary information R script

These will likely be calculated using your DPLYR skills, answering questions such as:

-   What is the average number of checkouts for each item?
-   What is the month or year with the most/least checkouts for a book that you're interested in?
-   What is the month or year with the most/least checkouts for ebooks?
-   How has the number of print book checkouts changed over time?

Feel free to calculate and report values that you find relevant.

Based on the data, we can see that Cain's book *Quiet: The Power of Introverts* has received noticeably more library checkouts from 2015 to 2023 compared to *Bittersweet: How Sorrow and Longing Make Us Whole*. 

### The Dataset

**Who collected/published the data? The data was collected and published**\
The date was collected by the *City of Seattle Open Data Portal* and is updated monthly.
This specific data set was collected from January 2012 to January 2023.
**What are the parameters of the data (dates, number of checkouts, kinds of books, etc.)?**\
The parameters of the data are Material Type, Title, Publisher, Number of Checkouts and Date, **How was the data collected or generated?**\
Based on the *City of Seattle Open Data Portal* FAQ, data comes from multiple current and historical sources.
Item checkouts of physical copies of books are sourced from the *Legrady artwork* archives while digital items are sourced from media vendors, which include *Overdrive, hoopla, Freegal and RBDigital*.
**Why was the data collected?**\
The data was collected to observe and compare trends between books written and published my author Susan Cain.
**What, if any, ethical questions do you need to consider when working with this data?**\
Will observing checkout data from these psychological books expose and stereotype certain individuals?
**What are possible limitations or problems with this data?(at least 200 words**\
A major limitation from this data set is that it only collects data from library checkouts that were done at a Seattle Public Library, which is one of the few libraries to make its circulation data public. It does not take into account data of private libraries, data in the United States or data on a more global scale. This data set also does not include data of purchases made at a bookstore, which is where a large number of people obtain books. This can alter the way the data is presented by not providing an accurate representation of trends, especially when it comes to measuring popularity and success of the author. In addition, *Quiet* was published in 2012 while *Bittersweet* was published in 2022. Since *Quiet* has existed longer than *Bittersweet*, it makes sense that the library checkouts will be significantly higher for *Quiet* so we can’t really conclude whether one book is better than the other. This data also cuts off on January 1, 2023, which can also be a huge limitation because *Bittersweet* was recently advertised by Oprah Winfrey in her March Book Club pick, so there is a possibility that there was a spike in library checkouts for this book this past month.

### First Trends Over Time Chart

Include a chart.
Make sure to describe why you included the chart, and what patterns emerged

The first chart that you will create and include will show the trend over time of your variable/topic/interest.
Think carefully about what you want to communicate to your user (you may have to find relevant trends in the dataset first!).
Here are some requirements to help guide your design:

-   Show more than one, but fewer than \~10 trends
    -   For example, two books, or 5 different mediums
-   You must have clear x and y axis labels
-   The chart needs a clear title 
-   You need a legend for your different line colors and a clear - legend title In your .Rmd file, make sure to describe why you included the chart, and what patterns emerged

When we say "clear" or "human readable" titles and labels, that means that you should not just display the variable name.

Here's an example of how to run an R script inside an RMarkdown file:

```{r Getting-Started-with-A3-SPL-Data.R, echo = FALSE)}
source('Getting-Started-with-A3-SPL-Data.R')
cain_plot
```

### Second Trends Over Time Chart

Include a chart.
Make sure to describe why you included the chart, and what patterns emerged

The second chart that you will create and include will show another trend over time of your variable/topic/interest.
Think carefully about what you want to communicate to your user (you may have to find relevant trends in the dataset first!).
Here are some requirements to help guide your design:

-   Show more than one, but fewer than \~10 trends
    -   For example, two books, or 5 different mediums
-   You must have clear x and y axis labels
-   The chart needs a clear title 
-   You need a legend for your different line colors and a clear - legend title In your .Rmd file, make sure to describe why you included the chart, and what patterns emerged

When we say "clear" or "human readable" titles and labels, that means that you should not just display the variable name.

Here's an example of how to run an R script inside an RMarkdown file:

\`\`\``{r, echo = FALSE, code = readLines("chart2_example.R")}`

### Line Plot of Library Checkouts

The last chart is up to you.
It could be a line plot, scatter plot, histogram, bar plot, stacked bar plot, and more.
Here are some requirements to help guide your design:

-   You must have clear x and y axis labels
-   The chart needs a clear title 
-   You need a legend for your different line colors and a clear legend title

Here's an example of how to run an R script inside an RMarkdown file:

```{r chart2_example.R, echo = FALSE}
source('chart2_example.R')
cain_plot_3
```
